Supporting Information: Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-05-25 Ben_Manuscripts/hdphmm/Supporting_Information.tex

@@

  \graphicspath{{./supporting_figures/}}
  \maketitle

@@

  The state sequence appears to converge within 500--1000 IHMM iterations. Step 1
  in the procedure from the previous section is the only time the state sequence 
  is determined. We ran 2000 iterations of the IHMM algorithm in order to arrive
  at a finalized state sequence. Stabilization of the VAR parameters indicates 
  that the state sequence has converged. In Figure~\ref{fig:convergence3d}, we plot
  the diagonal entries of 3 different $A$ and $\Sigma$ matrices as well as the 
  entries of the mean vector $\mathbf{c}$, fit to a single 3D methanol trajectory.
  States with a relatively small number of observations are more highly influenced 
  by the boundaries of the state segments and generally lead to much higher 
  variance of the $A$ and $\Sigma$ parameters. In these cases, the mean is
  a more reliable indicator of convergence since stabilized means imply that the
  fluctuations assigned to each state come from motion about a static location.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_47.pdf}
  \caption{706 total emissions}\label{fig:convergence3d_MET_low}
  \end{subfigure}
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_36.pdf}
  \caption{1000 total emissions}\label{fig:convergence3d_MET_medium}
  \end{subfigure}
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_59.pdf}
  \caption{201 total emissions}\label{fig:convergence3d_MET_high}
  \end{subfigure}
  \caption{Given a sufficient number of observations for a given mode, the VAR parameters
  appear to converge within 500 -- 1000 iterations. States with too few observations
  are more highly influenced by the boundaries of the state segments and generally
  lead to much higher fluctuations in the VAR parameters.}\label{fig:convergence3d}
  \end{figure}

  When we fix the state sequence in subsequent steps of the procedure, parameter 
  estimates converge within only a few steps. In Figure~\ref{fig:fixed_state_convergence}, we 
  plot the entries of $A$ and $\Sigma$ as a function of IHMM iteration. Some states
  are sampled far more frequently than others. High sampling frequently results in
  higher certainty converged parameter estimates. Due to fast convergence, we only
  run 100 iterations of the inference behavior when the state sequence is fixed.

@@

  \section{Demonstration of IHMM Parameterization Procedure}\label{section:ihmm_procedure}

  Parameterization of trajectories by the IHMM is a multi-stepped procedure. In the steps and
  figures that follow, we graphically illustrate the procedure which is described in detail 
  in Section~\ref{M-method:IHMM} of the main text.

  \begin{enumerate}
  	\item Parameterize in $x$, $y$, $z$ coordinates with $x$, $y$ coordinates relative to the nearest
  	pore center (see Figure~\ref{fig:xyz_hmm}). 		
  	\item Cluster the VAR parameters from the states found in all 24 trajectories (See 
  	Section~\ref{method:clustering} of the main text). Reassign the state sequence so that segments
  	which belong to the same cluster are labeled the same across all solute trajectories.
  	\item Zero the trajectories. First, zero out the $y$ dimension by rotating each segment of 
  	the trajectory, as partitioned before clustering, about the $z$ axis, in order to align the 
  	mean $xy$ vector with the $x$ axis (see Figure~\ref{fig:y_zeroed}). Then, subtract the mean in
  	$x$ and $z$ (Figure~\ref{fig:xyz_zeroed}.
  	\item Fix clustered state sequence and infer parameters of each state, assuming a mean of zero
  	for all states (see Figure~\ref{fig:zeroed_clustered_hmm}). 
  \end{enumerate}

  \begin{figure}
    \centering
	\begin{subfigure}{0.6\textwidth}
	\includegraphics[width=\textwidth]{xyz_hmm.pdf}
		\caption{}\label{fig:xyz_hmm}
	\end{subfigure}  
		\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{y_zeroed.pdf}
		\caption{}\label{fig:y_zeroed}
	\end{subfigure}  	
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{xyz_zeroed.pdf}
		\caption{}\label{fig:xyz_zeroed}
	\end{subfigure}  		
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{zeroed_clustered_hmm.pdf}
		\caption{}\label{fig:zeroed_clustered_hmm}
	\end{subfigure}  
	\caption{IHMM parameterization is a multi-step procedure. In the above plots, distinct colors
	correspond to distinct state behavior. (a) First we parameterize the trajectories in terms of
	the $x$, $y$, $z$ solute center-of-mass coordinates with the $x$ and $y$ coordinates relative
	to the nearest pore center (b) Next, we convert the $x$ and $y$ 	
	coordinates to radial coordinates, $r$, then fix the state sequence and use the inference 
	component of the IHMM in order to estimate the parameters in terms of $r$ and $z$. Note that
	the segment colors in (b) are the same as (a) because we fixed the state sequence.
	(c) After clustering the VAR parameters and reassigning the state sequence so that segments 
	which belong to the same cluster are labeled the same across all solute trajectories, we subtract
	the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory, and then fix the clustered state sequence to infer the parameters of each clustered 
  	state, assuming a mean of zero for all states. Note that the colors are no longer aligned with 
  	figures (a) and (b) because the states are relabeled in terms of their clusters. There are
  	also less states in this single trajectory.
	}\label{fig:hmm_demo}
  \end{figure}

@@

  \textit{Choosing Prior Parameters}: The states identified by the IHMM are 
  heavily influenced by the Gaussian prior placed on $\mathbf{c}$ in 
  Equation~\ref{M-eqn:var} of the main text. The entries of $A$ and $\Sigma$
  do not vary over a wide range, so the final parameters were relatively 
  insensitive to the priors. In order to maximally automate the IHMM procedure,
  we attempted to parameterize the
  prior on $\mathbf{c}$ in an intelligent way. The prior parameters should
  be chosen such that the mean level of each state lies within a region of 
  reasonable probability of the prior (see Figure~\ref{fig:prior_guesses}).
  In each dimension, we defined the prior mean to lie halfway between the 
  maximum and minimum of each trajectory dimension. To parameterize the 
  prior's variance in each dimension, we defined the maximum and minimum to 
  be 2 standard deviations from the prior mean. Although this approach has 
  worked quite well for the data in this work, it is important to check the 
  results to determine whether further adjustments to the prior might be needed.
%  In Figure TBD, we show the result of a parameterization where the prior
%  parameters of $c$ were poorly chosen. %TODO

@@

  \pagebreak

  The best way to ensure reliable MSD predictions is to ensure proper reproduction
  of the individual MD MSDs by realizations of the IHMM. Over-predictions and 
  under-predictions are usually the result of a poorly identified state sequence. (see
  Figure TBD for example)

  \begin{figure}
  \centering
  \begin{subfigure}{0.35\textwidth}
  \includegraphics[width=\textwidth]{overestimate_ACH_21.pdf}
  \caption{}\label{fig:msd_overestimate}
  \end{subfigure}
  \begin{subfigure}{0.63\textwidth}
  \includegraphics[width=\textwidth]{state_sequence_before_ACH_21.pdf}
  \caption{}\label{fig:state_sequence_before}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  \includegraphics[width=\textwidth]{msd_improvement_ACH_21.pdf}
  \caption{}\label{fig:msd_improvement}
  \end{subfigure}
  \begin{subfigure}{0.63\textwidth}
  \includegraphics[width=\textwidth]{state_sequence_after_ACH_21.pdf}
  \caption{}\label{fig:state_sequence_after}
  \end{subfigure}
  \caption{(a, b) In some cases, the state segmentation leads to over- or 
  under-prediction of the MSD, usually due to poor initialization of the state
  sequence. In (b), state transitions are too frequent and lead to an
  overprediction of the MD MSD. (c) We can greatly improve MSD predictions by
  giving the IHMM a good guess at the initial state sequence. (d) A better
  initial state segmentation via seeding, gives more reasonable estimates
  of the final state sequence.}\label{fig:improvement_state_sequence}
  \end{figure}

  There are two primary approaches we have implemented in order to improve the final 
  state sequence predictions. Both attempt to aid the initial state assignments 
  algorithmically. First, we always initialize the prior on $\mathbf{c}$ in 
  Equation~\ref{M-eqn:var} of the main text as described in 
  Section~\ref{section:prior_guesses}. In particularly tricky cases, the initial
  state sequence estimate is so poor that qualitative trajectory realizations
  and MSD predictions are irreconcilable with MD. In our second approach, we can
  seed the initial state sequence with a guess made by applying the IHMM to 
  smaller segments of the trajectory.

  \begin{figure}
  \centering
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment0.pdf}
  \caption{Found 4 Unique States}\label{fig:segment1}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment1.pdf}
  \caption{Found 14 Unique States}\label{fig:segment2}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment2.pdf}
  \caption{Found 13 Unique States}\label{fig:segment3}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment3.pdf}
  \caption{Found 12 Unique States}\label{fig:segment4}
  \end{subfigure}
  \begin{subfigure}{1\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_full.pdf}
  \caption{Initial state sequence seeded with states found in (a)--(d)}\label{fig:full}
  \end{subfigure}
  \caption{In (a--d), we show the state sequence resulting from 5 iterations
  of the IHMM on each of the four quarters of the center-of-mass trajectory in
  (e) after splitting it into 4 separate segments. We use the concatenation 
  of the four state sequences as the initial sequence for the IHMM run on the 
  fully trajectory in (e).
  }\label{fig:seed_sequence}
  \end{figure}

@@

  % BJC: will show Gaussian mixture model which has a cluster engulfing another cluster, so
  % it's tails are on both sides of the center cluster.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.95\textwidth}
  \includegraphics[width=\textwidth]{bayesian_A.pdf}
  \caption{Bayesian Gaussian mixture model}\label{fig:bayesian_A}
  \end{subfigure}
  \begin{subfigure}{0.95\textwidth}
  \includegraphics[width=\textwidth]{agglomerative_A.pdf}
  \caption{Agglomerative Clustering}\label{fig:agglomerative_A}
  \end{subfigure}
  \caption{To reduce the state space, we prefer agglomerative clustering over Bayesian
  Gaussian mixture modeling. In the plots above we show the results of clustering the
  diagonal entries of the autoregressive coefficient matrices, $A$, of methanol. (a) 
  Bayesian Gaussian mixture models tend to delocalize the clusters in parameter space
  since the Gaussians can overlap. Agglomerative clustering prevents overlap of 
  clusters and ensures that all states within each cluster have similar parameters.
  }\label{fig:clustering_choice}
  \end{figure}

@@

%  \subsection{Clustering on the self-transition probabilities}\label{section:selfT_cluster}
%  
%  To cluster on $T_{ii}$, we first cast it in terms of the expected dwell times. The expected 
%  value of the number of sequential self-transitions is simply $\frac{1}{1 - T_{ii}}$. This 
%  relationship correctly implies that dwell times approach infinity as $T_{ii}$ approaches 1. 
%  There we found the most success with agglomerative clustering by linearizing this 
%  relationship and clustering on $-log(1 - T_{ii})$.

@@

  In this section, we determine the optimal linkage criteria and number of clusters
  for studying methanol's parameters. We applied the same workflow to make analogous
  decisions for ethylene glycol, urea and acetic acid. In all cases, we found that
  the `ward' linkage criteria resulted in the most useful clustering. We chose to 
  use x number of clusters for each solute, the minimum number of clusters, by our
  analysis, needed to adequately distinguish state dynamics.

  In order to cluster the parameter sets, we need to determine the linkage criteria 
  that will be used to measure the distance between clusters as well as an appropriate
  total number of clusters. In Figure~\ref{fig:linkages}, we show the result of clustering
  with each type of linkage criteria available with the 
  \texttt{sklearn.cluster.AgglomerativeClustering} class of the \texttt{scikit-learn} 
  python package. It is clear that both the `average' and `single' linkage criteria 
  do not sufficiently cluster the data. The `ward' criteria appears to result in the 
  highest number of large population clusters.

@@

  We use the Silhouette score in order to aid us in choosing the total number of
  clusters. In Figure~\ref{fig:silhouette_MET}, we plot the Silhouette score as
  a function of the total number of clusters using the `ward' linkage criteria. 
  A Silhouette score of 1 indicates the best clustering possible, values near
  0 indicate overlapping clusters, and negative values indicate samples belong to
  the wrong clusters. For the methanol parameters, the Silhouette score drops off
  quickly with number of clusters before plateauing around 7 total clusters. 
  However, given the low spread in $\mu_r$, it is not clear that using 7 or less 
  clusters will give useful clustering results.

@@

  \begin{figure}
  \begin{subfigure}{0.9\textwidth}
  \includegraphics[width=1\textwidth]{clustered_traj_MET_ward_4.pdf}
  \caption{4 total clusters}\label{fig:4cluster_state_sequence}
  \end{subfigure}
  \begin{subfigure}{0.9\textwidth}
  \includegraphics[width=1\textwidth]{clustered_traj_MET_ward_12.pdf}
  \caption{12 total clusters}\label{fig:12cluster_state_sequence}
  \end{subfigure}
  \begin{subfigure}{0.9\textwidth}
  \includegraphics[width=1\textwidth]{clustered_traj_MET_ward_20.pdf}
  \caption{20 total clusters}\label{fig:20cluster_state_sequence}
  \end{subfigure}
  \caption{Although the Silhouette score for four total clusters is high, we do
  not see adequate distinction between clusters until we group the parameters
  from the 24 trajectories into 20 total clusters. When we use 12 total clusters,
  the distinction appears acceptable except that the segment from ~50--550 ns
  and from 1200--1500 ns (both colored orange) are clustered together. The radial
  means of these two segments correspond to two very different morphological
  regions of the membrane.}\label{fig:clustered_state_sequences}
  \end{figure}

  \begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{nclusters_msd_MET_ward.pdf}
  \caption{The total MSD predicted by stochastic realizations of the clustered
  model is relatively insensitive to the total number of clusters. All
  predictions are less than the predictions of MD and the unclustered trajectories.
  This behavior is expected due to underestimation of the short dwell time 
  densities as described in Section~\ref{M-section:unclustered_MSD_prediction}
  of the main text.}\label{fig:MET_nclusters} 
  \end{figure}

  This clustering analysis will be finished after some mechanistic studies. It is
  clear that we will need some understanding of mechanisms in order to 
  arrive at a cluster recommendation.

%  \begin{figure}[h]
%  \centering
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{nclusters_msd_MET_ward.pdf}
%  \caption{ward}\label{fig:nclusters_msd_MET_ward}
%  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{nclusters_msd_MET_complete.pdf}
%  \caption{complete}\label{fig:nclusters_msd_MET_complete}
%  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{nclusters_msd_MET_average.pdf}
%  \caption{average}\label{fig:nclusters_msd_MET_average}
%  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{nclusters_msd_MET_single.pdf}
%  \caption{single}\label{fig:nclusters_msd_MET_single}
%  \end{subfigure} 
%  \caption{}\label{fig:MET_nclusters} 
%  \end{figure}

  \section{Deviations of the IHMM from Molecular Motion}\label{section:shorttimes_msd}

   At extremely short time lags, the MSD is generally over-predicted by the IHMM
  (see Figure~\ref{fig:unclustered_msd_shortlag}) because the VAR(1) model assumes
  multivariate Gaussian noise while the MD data suggests the noise is better 
  modeled by more general L\'evy stable noise. The first time lag of the MSD is
  equivalent to the variance of all observed fluctuations (first-order difference). 
  The IHMM simulation appears to over-estimate the variance because the actual 
  distribution of solute fluctuations has heavy tails with a smaller variance (Figure~\ref{fig:emission_widths}).
  The total distribution of hops is expected to be heavy tailed because it is a
  combination of hop distributions from many states with different variances. More
  interestingly, in Figure ~\ref{fig:state_emission_widths}, we show that the 
  emission distribution of a single state picked out by the IHMM is fit better by
  a more general L\'evy distribution while the emissions from IHMM realizations are
  Gaussian, as expected. An improvement to the IHMM that may allow dynamics more 
  faithful to MD would use an autoregressive model with L\'evy stable noise. However,
  this would be less computationally efficient due to lack of conjugate priors.

  \begin{figure}
  \centering
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{unclustered_msd_MET_shortlag.pdf}
  \caption{}\label{fig:unclustered_msd_shortlag}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{emission_widths.pdf}
  \caption{}\label{fig:emission_widths}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{state_emission_widths.pdf}
  \caption{}\label{fig:state_emission_widths}
  \end{subfigure}
  \caption{(a) At short time lags, the MD MSD (black) and the IHMM MSD (orange)
  differ significantly in their shape and curvature. (b) While the 
  fluctuations in $z$ of both distributions are heavy-tailed, fitting well to zero-centered
  symmetric L\'evy stable distributions with stability parameters, $\alpha$, the IHMM has
  a wider variance. (c) The distribution of fluctuations in a chosen state frequently
  visited by methanol is heavy tailed, but the IHMM is constrained to produce fluctuations
  from a multivariate normal distribution ($\alpha$=2 in each dimension).
  }\label{fig:short_timelags}
  \end{figure}

  Also apparent from Figure~\ref{fig:unclustered_msd_shortlag} is that the 
  curvature of the MD MSD is far more gradual than that generated by the IHMM.
  This is because the VAR(1) model is only correlated to its previous fluctuation
  while the MD simulations suggest that correlation persists for at least tens
  of nanoseconds.
    % BJC2: could probably use a correlation function to suggest the AR order and stick the plot in SI
  One may be able to reproduce the curvature at short time lags with higher order
  VAR models, but it becomes a much higher dimensional parameterization that would
  require significantly more data in order to produce stable results. Since we 
  are interested in reproducing longer time scale behavior, it is not an issue of
  great concern.

  \section{Influence of number of clusters on qualitative hybrid trajectories}

>>>>

\usepackage{tocloft}

\renewcommand\cftsecfont{\normalfont}
\renewcommand\cftsecpagefont{\normalfont}
\renewcommand{\cftsecleader}{\cftdotfill{\cftsecdotsep}}
\renewcommand\cftsecdotsep{\cftdot}
\renewcommand\cftsubsecdotsep{\cftdot}
\cftsetindents{section}{0em}{3em}
%\cftsetindents{subsection}{1em}{3em}

@@

  \graphicspath{{./supporting_figures/}}
  \maketitle
  \tableofcontents

@@

  The state sequence appears to converge within 500--1000 IHMM iterations. To
  be sure, we ran 2000 iterations of the IHMM algorithm in order to arrive at a 
  finalized state sequence. Stabilization of the VAR parameters indicates 
  that the state sequence has converged. In Figure~\ref{fig:convergence3d}, we plot
  the diagonal entries of the $A$ and $\Sigma$ matrices as well as the entries of
  the mean vector $\mathbf{c}$, for three different states identified in a 3D 
  methanol trajectory. States with a relatively small number of observations are
  more highly influenced by the boundaries of the state segments and generally 
  lead to much higher variance of the $A$ and $\Sigma$ parameters. In these cases,
  the mean is a more reliable indicator of convergence since stabilized means 
  imply that the fluctuations assigned to each state come from motion about a 
  static location.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_47.pdf}
  \caption{706 total emissions}\label{fig:convergence3d_MET_low}
  \end{subfigure}
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_36.pdf}
  \caption{1000 total emissions}\label{fig:convergence3d_MET_medium}
  \end{subfigure}
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_59.pdf}
  \caption{201 total emissions}\label{fig:convergence3d_MET_high}
  \end{subfigure}
  \caption{Given a sufficient number of observations for a given mode, the VAR parameters
  appear to converge within 500 -- 1000 iterations. States with fewer observations
  are more highly influenced by the boundaries of the state segments which can
  lead to much higher fluctuations in the VAR parameters.}\label{fig:convergence3d}
  \end{figure}

  When we fix the state sequence during the steps of the procedure following the 
  initial state sequence determination, the parameter estimates converge within 
  only a few steps. In Figure~\ref{fig:fixed_state_convergence}, we plot the 
  entries of $A$ and $\Sigma$ as a function of IHMM iteration. Some states are 
  sampled far more frequently than others. High sampling frequently results in 
  higher certainty converged parameter estimates. Due to fast convergence, we 
  only ran 100 iterations of the inference behavior when the state sequence is fixed.

@@

@@

  \textit{Choosing Prior Parameters}: The states identified by the IHMM are 
  heavily influenced by the Gaussian prior placed on $\mathbf{c}$ in 
  Equation~\ref{M-eqn:var} of the main text. The entries of $A$ and $\Sigma$
  do not vary over a wide range, so the final parameters were relatively 
  insensitive to the priors. In order to maximally automate the IHMM procedure,
  we attempted to parameterize the
  prior on $\mathbf{c}$ in an intelligent way. The prior parameters should
  be chosen such that the mean level of each state lies within a region of 
  reasonable probability of the prior (see Figure~\ref{fig:prior_guesses}).
  In each dimension, we defined the prior mean to lie halfway between the 
  maximum and minimum of each trajectory dimension. To parameterize the 
  prior's variance in each dimension, we defined the maximum and minimum to 
  be 2 standard deviations from the prior mean. Although this approach has 
  worked quite well for the data in this work, it is important to check the 
  results to determine whether further adjustments to the prior might be needed.

@@

  \textit{Seeding the initial state sequence}: We always initialize the prior on
  $\mathbf{c}$ as described above but in some 
  particularly tricky cases, the initial state sequence estimate is so poor that
  qualitative trajectory realizations and MSD predictions are irreconcilable with
  MD. A simple way to identify bad parameterizations is to check whether the MD 
  MSD of individual trajectories can be reasonably reproduced by realizations of
  the HDP-AR-HMM fit to the same trajectory. In Figure~\ref{fig:msd_overestimate},
  the MSD predicted by the HDP-AR-HMM dwarfs that calculated from MD, due to 
  rapid switching between states in the final state sequence.

  We can improve the final state sequence estimate by seeding the initial state 
  sequence with a guess made by applying the HDP-AR-HMM to smaller segments of the 
  trajectory. In Figure~\ref{fig:seed_sequence}, we illustrate this process. In
  Figure~\ref{fig:msd_improvement}, the MSD predicted by the HDP-AR-HMM is
  much more consistent with MD after applying this approach.

  \begin{figure}
  \centering
  \begin{subfigure}{0.35\textwidth}
  \includegraphics[width=\textwidth]{overestimate_ACH_21.pdf}
  \caption{}\label{fig:msd_overestimate}
  \end{subfigure}
  \begin{subfigure}{0.63\textwidth}
  \includegraphics[width=\textwidth]{state_sequence_before_ACH_21.pdf}
  \caption{}\label{fig:state_sequence_before}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  \includegraphics[width=\textwidth]{msd_improvement_ACH_21.pdf}
  \caption{}\label{fig:msd_improvement}
  \end{subfigure}
  \begin{subfigure}{0.63\textwidth}
  \includegraphics[width=\textwidth]{state_sequence_after_ACH_21.pdf}
  \caption{}\label{fig:state_sequence_after}
  \end{subfigure}
  \caption{(a, b) In some cases, the state segmentation leads to over- or 
  under-prediction of the MSD, usually due to poor initialization of the state
  sequence. In (b), state transitions are too frequent, particularly between 
  the light blue and light green states, and ultimately causes an over-prediction
  of the MD MSD. (c) We can greatly improve MSD predictions by giving the HDP-AR-HMM a good
  guess at the initial state sequence. (d) A better initial state segmentation
  via seeding, gives more reasonable estimates of the final state sequence
  as illustrated in Figure~\ref{fig:seed_sequence}.}\label{fig:improvement_state_sequence}
  \end{figure}

  \begin{figure}
  \centering
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment0.pdf}
  \caption{Found 4 Unique States}\label{fig:segment1}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment1.pdf}
  \caption{Found 14 Unique States}\label{fig:segment2}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment2.pdf}
  \caption{Found 13 Unique States}\label{fig:segment3}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_segment3.pdf}
  \caption{Found 12 Unique States}\label{fig:segment4}
  \end{subfigure}
  \begin{subfigure}{1\textwidth}
  \includegraphics[width=\textwidth]{seed_ACH21_full.pdf}
  \caption{Final state sequence seeded with states found in (a)--(d)}\label{fig:full}
  \end{subfigure}
  \caption{In (a--d), we show the state sequences which result from 5 iterations
  of the HDP-AR-HMM on each of the four quarters of the center-of-mass trajectory
  in (e). We use the concatenation of the four state sequences as the initial 
  sequence for the HDP-AR-HMM inference run on the fully trajectory in (e).
  }\label{fig:seed_sequence}
  \end{figure}

@@

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.95\textwidth}
  \includegraphics[width=\textwidth]{bayesian_A.pdf}
  \caption{Dirichlet process Gaussian mixture model}\label{fig:bayesian_A}
  \end{subfigure}
  \begin{subfigure}{0.95\textwidth}
  \includegraphics[width=\textwidth]{agglomerative_A.pdf}
  \caption{Agglomerative Clustering}\label{fig:agglomerative_A}
  \end{subfigure}
  \caption{To reduce the state space, we prefer agglomerative clustering over non-parametric
  Dirichlet process Gaussian mixture modeling. In the plots above we show the results of clustering the
  diagonal entries of the autoregressive coefficient matrices, $A$, of methanol. (a) 
  Dirichlet process Gaussian mixture models tend to delocalize the clusters in parameter space
  since the Gaussians can overlap. Agglomerative clustering prevents overlap of 
  clusters and ensures that all states within each cluster have similar parameters.
  }\label{fig:clustering_choice}
  \end{figure}

@@

@@

  In this section, we determine the optimal linkage criteria and number of clusters
  for studying methanol's parameters. We applied the same workflow in order to make
  analogous decisions for ethylene glycol, urea and acetic acid. In all cases, we 
  found that the `ward' linkage criteria resulted in the most useful clustering. 
  We chose to use 10 clusters for each solute, the minimum number of clusters, by our
  analysis, needed to adequately distinguish state dynamics.

  In order to cluster the parameter sets, we need to determine the linkage criteria 
  that will be used to measure the distance between clusters as well as an appropriate
  total number of clusters. In Figure~\ref{fig:linkages}, we show the result of clustering
  with each type of linkage criteria available with the \\
  \texttt{sklearn.cluster.AgglomerativeClustering} class of the \texttt{scikit-learn} 
  python package. It is clear that both the `average' and `single' linkage criteria 
  do not sufficiently cluster the data. The `ward' criteria appears to result in the 
  highest number of large population clusters.

@@

  We use the Silhouette score in order to aid us in choosing the total number of
  clusters. In Figure~\ref{fig:silhouette_MET}, we plot the Silhouette score as
  a function of the total number of clusters using the `ward' linkage criteria. 
  A Silhouette score of 1 indicates the best clustering possible, values near
  0 indicate overlapping clusters, and negative values indicate that samples belong to
  the wrong clusters. For the methanol parameters, the Silhouette score drops off
  quickly with number of clusters before plateauing around 7 total clusters. 
  However, given the low spread in $\mu_r$, it is not clear that using 7 or less 
  clusters will give useful clustering results.

@@

  \begin{figure}
  \begin{subfigure}{0.9\textwidth}
  \includegraphics[width=1\textwidth]{clustered_traj_MET_ward_4.pdf}
  \caption{4 total clusters}\label{fig:4cluster_state_sequence}
  \end{subfigure}
  \begin{subfigure}{0.9\textwidth}
  \includegraphics[width=1\textwidth]{clustered_traj_MET_ward_10_nolabel.pdf}
  \caption{10 total clusters}\label{fig:20cluster_state_sequence}
  \end{subfigure}
  \caption{Although the Silhouette score for four total clusters is high, we do
  not see adequate distinction between clusters until we group the parameters
  from the 24 trajectories into at least 10 total clusters.}\label{fig:clustered_state_sequences}
  \end{figure}

  \section{Obtaining clustered state parameters}\label{section:ihmm_procedure}

  Parameterization the clustered states with the HDP-AR-HMM is a multi-stepped procedure. 
  In the steps and figures that follow, we graphically illustrate the procedure which is
  described with additional detail in Section~\ref{M-method:clustering} of the main text.

  \begin{enumerate}
  	\item Parameterize in $x$, $y$, $z$ coordinates with $x$, $y$ coordinates relative to the nearest
  	pore center (see Figure~\ref{fig:xyz_hmm}). 		
  	\item Cluster the VAR parameters from the states found in all 24 trajectories (See 
  	Section~\ref{M-method:clustering} of the main text). Reassign the state sequence so that segments
  	which belong to the same cluster are labeled the same across all solute trajectories.
  	\item Zero the trajectories. First, zero out the $y$ dimension by rotating each segment of 
  	the trajectory, as partitioned before clustering, about the $z$ axis, in order to align the 
  	mean $xy$ vector with the $x$ axis (see Figure~\ref{fig:y_zeroed}). Then, subtract the mean in
  	$x$ and $z$ (Figure~\ref{fig:xyz_zeroed}).
  	\item Fix clustered state sequence and infer parameters of each state, assuming a mean of zero
  	for all states (see Figure~\ref{fig:zeroed_clustered_hmm}). 
  \end{enumerate}

  \begin{figure}
    \centering
	\begin{subfigure}{0.6\textwidth}
	\includegraphics[width=\textwidth]{xyz_hmm.pdf}
		\caption{}\label{fig:xyz_hmm}
	\end{subfigure}  
		\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{y_zeroed.pdf}
		\caption{}\label{fig:y_zeroed}
	\end{subfigure}  	
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{xyz_zeroed.pdf}
		\caption{}\label{fig:xyz_zeroed}
	\end{subfigure}  		
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[width=\textwidth]{zeroed_clustered_hmm.pdf}
		\caption{}\label{fig:zeroed_clustered_hmm}
	\end{subfigure}  
	\caption{In the above plots, distinct colors correspond to distinct state behavior. 
	(a) First we parameterize the trajectories in terms of the $x$, $y$, $z$ solute 
	center-of-mass coordinates with the $x$ and $y$ coordinates relative to the nearest 
	pore center. The parameters of these states are used to predict the unclustered MSDs
	according to method 1 described in Section~\ref{M-method:realizations} of the main
	text. (b) Next, we zeroed out the $y$ dimension by rotating each segment of the 
	trajectory, as partitioned before clustering, about the $z$ axis, in order to align the 
  	mean $xy$ vector with the $x$ axis. (c) We then subtracted the $x$ and $y$ means of
  	each segment in order to fully zero the trajectory. (d) Finally, we applied the 
  	inference component of the HDP-AR-HMM in order to infer the parameters of the 
  	clustered states.
	}\label{fig:hmm_demo}
  \end{figure}

  \section{Cause of underestimate of urea's MD MSD}\label{section:urea_underestimate}

  Realizations of the HDP-AR-HMM underestimate the MD MSD of urea due to MD
  trajectories with axial motion correlation that is likely uncharacteristic and
  unable to be captured by ensembles of stochastic HDP-AR-HMM realizations. One
  such trajectory is plotted in Figure~\ref{fig:urea_underestimate}. Axial hops 
  appear positively correlated, jumping consistently in the negative $z$ direction, 
  leading to a high MSD. The parabolic trajectory of the MD MSD curve implies
  underlying superdiffusion which would require some kind of facilitating
  driving force. It is likely that running the trajectory longer would reduce
  the slope of the MSD curve, making it more consistent with the HDP-AR-HMM 
  prediction.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.35\textwidth}
  \includegraphics[width=\textwidth]{underestimate_URE_13.pdf}
  \caption{}\label{fig:underestimate_msd}
  \end{subfigure}
  \begin{subfigure}{0.63\textwidth}
  \includegraphics[width=\textwidth]{state_sequence_before_URE_13.pdf}
  \caption{}\label{fig:underestimate_traj}
  \end{subfigure}
  \caption{(a) The HDP-AR-HMM under-predicts the MD MSD of urea when fit
  to the trajectory in (b) because axial hops are highly correlated, 
  behavior our model is unable to reproduce.}\label{fig:urea_underestimate}
  \end{figure}

  \section{Influence of number of clusters on qualitative hybrid trajectories}\label{section:qualitative}

  \begin{figure}[h]
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{qualitative_clustered_MET_20.pdf}
  \caption{20 clusters}\label{fig:qualitative_20}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{qualitative_clustered_MET_30.pdf}
  \caption{30 clusters}\label{fig:qualitative_30}
  \end{subfigure}
  \caption{The quality of hybrid trajectories resulting from clustered parameters
  sets qualitatively improves as the number of clusters is increased. This comes 
  at the cost of a larger state space to interpret.}\label{fig:qualitative_improvement}
  \end{figure}

  \section{Deviations of the HDP-AR-HMM from Molecular Motion}\label{section:shorttimes_msd}
